\section{Results and Discussion}\label{ch:tmf:exp_res}


We evaluated the performance of MF method and the proposed methods, i.e.,
\TMF (TMF) and \TMF with Dropout (TMF + Dropout) on the real datasets presented in Table~\ref{table:ch:matcomp:datasets_table}. 


\subsection{Performance for rating prediction on real datasets}
We investigated the performance of the different methods for the task
of rating predictions on different rating datasets. Table~\ref{table:drop_test_rmse} shows the best performance
achieved by the different methods on the datasets for different ranks. As can be seen in the table,
the proposed approaches outperform the MF method for the task of rating
predictions in different datasets. The performance is significantly better for
FX and EM datasets. 
The better performance of the proposed methods for FX and EM datasets is because
in these datasets there exists either the users or the items that have few
ratings and the proposed approaches are effective in selecting the
few ranks that can be estimated accurately for these users or
items.
Also, for NF dataset which has a high average number of ratings for the users
and the items the performance of the proposed approaches is similar to that of
the MF method. This is because the proposed methods use all the available ranks
as active ranks in such a
dataset and hence essentially reduces to MF method.


Additionally, among the proposed methods, TMF +  Dropout
outperforms the TMF, and this improved performance of TMF + Dropout illustrates
its effectiveness in preventing overfitting and generating better predictions compared to that of the TMF
method. 
Furthermore, TMF + Dropout needs fewer ranks in comparison to that of MF to
achieve the same performance for the majority of the datasets thereby
illustrating its advantage over MF in terms of memory and storage requirements.


\subsection{Performance for the users and the items with different number of
ratings}
We investigated how does the performance of the MF and the proposed methods vary
for the users and the items with the different number of ratings.
Table~\ref{table:drop_test_quart_rmse} shows
the performance of the different methods for different quartiles of the users
and the items ordered by their frequency. As can be seen in the table, the
proposed methods outperform the MF method for lower quartiles for
all the datasets and to some extent performs better for higher quartiles as
well.
The performance of the proposed method is significantly better for lower
quartiles in FX, EM and to some extent in ML datasets, as these datasets contain users or
items with few ratings.
The better performance of the proposed methods for the users and the items 
with few ratings is because unlike MF the proposed approaches consider only
a subset of ranks that are estimated accurately while trying to predict
ratings that involve infrequent users or items.
Also, since NF dataset has a high average number of ratings, the proposed
approaches use all the ranks as active ranks thereby giving similar performance 
as that of the MF method for all the quartiles.


Also, similar to our results for overall rating prediction the  TMF +
Dropout method outperforms the TMF method for different quartiles for most of the
datasets thereby demonstrating its advantage over the TMF method.


\begin{table*}[hbt]
\caption{Test RMSE of the proposed approaches for different datasets. We also
  show the RMSE for the users and the items in
different quartiles created in increasing order by their frequency. Q1 refers to
the quartile containing the least frequent users or items followed by remaining
in Q2, Q3, and
Q4.
}
\label{table:drop_test_quart_rmse}
  
  \begin{center}
    \begin{tabular}{lrrr|rrr}
       & \multicolumn{3}{c}{EM} & \multicolumn{3}{c}{FX} \\
      \hline
      & \multicolumn{1}{p{1.7cm}}{\centering MF \\ (Rank 50)} &
      \multicolumn{1}{p{1.7cm}}{\centering TMF \\ (Rank 50)} &
     \multicolumn{1}{p{1.7cm}}{\centering TMF \\ + Dropout \\ (Rank 10)} &  
      \multicolumn{1}{p{1.7cm}}{\centering MF \\ (Rank 10)} &
      \multicolumn{1}{p{1.7cm}}{\centering TMF \\ (Rank 10)}
      & \multicolumn{1}{p{1.7cm}}{\centering TMF \\ + Dropout \\ (Rank 15)} \\  
      \hline
      Overall & 1.166 & 1.161 & \underline{1.155} & 0.868 & 0.851 &
      \underline{0.847} \\
      \hline
      Item Q1 & 1.626 & 1.584 & \underline{1.258} & 1.607  & \underline{1.252} & 1.256 \\ 
      Item Q2 & 1.177 & 1.165 & \underline{1.139} & 1.190  & \underline{0.944} & \underline{0.944} \\ 
      Item Q3 & 1.106 & 1.107 & \underline{1.104} & 0.898  & 0.785 &
      \underline{0.780} \\ 
      Item Q4 & 1.170 & 1.165 & \underline{1.161} & 0.864  & 0.851 &
      \underline{0.847} \\ 
      \hline
      User Q1 & 1.600 & 1.437 & \underline{1.373} & 1.478 & \underline{1.247} &
      1.260 \\ 
      User Q2 & 1.242 & 1.240 & \underline{1.234} & 1.318 & \underline{1.144} &
      1.151 \\ 
      User Q3 & \underline{1.180} & 1.192 & 1.192 & 1.053 & \underline{0.964} &
      \underline{0.964} \\ 
      User Q4 & 1.119 & 1.119 & \underline{1.115} & 0.850 & 0.841 &
      \underline{0.836} \\
      \hline
    \end{tabular}
  \end{center}
  
  \begin{center}
    \begin{tabular}{lrrr|rrr}
      & \multicolumn{3}{c}{ML} & \multicolumn{3}{c}{NF} \\
      \hline
      & \multicolumn{1}{p{1.7cm}}{\centering MF \\ (Rank 25)} &
      \multicolumn{1}{p{1.7cm}}{\centering TMF \\ (Rank 50)} &
      \multicolumn{1}{p{1.7cm}}{\centering TMF\\ + Dropout \\ (Rank 25)} 
      & \multicolumn{1}{p{1.7cm}}{\centering MF \\ (Rank 50)} &
      \multicolumn{1}{p{1.7cm}}{\centering TMF \\ (Rank 50)} &
      \multicolumn{1}{p{1.7cm}}{\centering TMF\\ + Dropout \\ (Rank 50)} 
      \\  
      \hline
      Overall & 0.809 & 0.806 &  \underline{0.804} & 0.830 & \underline{0.829} &
      \underline{0.829}\\
      \hline
      Item Q1 & 2.347 & 2.390 & \underline{2.115} & 0.958 & 0.959 &
      \underline{0.956} \\ 
      Item Q2 & 1.396 & 1.435 & \underline{1.123} & 0.935 & 0.935 &
      \underline{0.932} \\ 
      Item Q3 & 0.867 & 0.874 & \underline{0.851} & 0.887 & 0.887 &
      \underline{0.884} \\ 
      Item Q4 & 0.804 & \underline{0.801} & \underline{0.801} & 0.824 &
\underline{0.823} & \underline{0.823} \\ 
      \hline
      User Q1 & 1.130 & 1.119 & \underline{1.078} & 1.019 & 1.019 &
      \underline{1.014} \\ 
      User Q2 & \underline{0.967} & 0.969 & 0.968 & 0.923 & 0.923 &
      \underline{0.921} \\ 
      User Q3 & \underline{0.856} & 0.859 & 0.863 & 0.859 & 0.859 &
      \underline{0.858} \\ 
      User Q4 & 0.774 & 0.770 & \underline{0.769} & 0.803 & \underline{0.802} &
      \underline{0.802}\\
      \hline
    \end{tabular}
  \end{center}

  \iffalse
    \begin{center}
      \begin{tabular}{lrrr}
         & \multicolumn{3}{c}{BX }  \\
        \hline
        & \multicolumn{1}{p{1.5cm}}{\centering MF \\ (Rank 1)} &
        \multicolumn{1}{p{1.5cm}}{\centering TMF \\ (Rank 5)} &
        \multicolumn{1}{p{1.5cm}}{\centering TMF\\ + Dropout \\ (Rank 50)} 
        \\  
        \hline
        Overall & \underline{3.609} & 3.617 & 3.611  \\
        \hline
        Item Q1 & 3.861 & 3.834 & \underline{3.831}   \\ 
        Item Q2 & 3.696 & \underline{3.641} & \underline{3.641}  \\ 
        Item Q3 & 3.600 & 3.567 & \underline{3.565}   \\ 
        Item Q4 & \underline{3.542} & 3.578 & 3.569   \\ 
        \hline
        User Q1 & 4.410 & \underline{4.351} & 4.361   \\ 
        User Q2 & \underline{4.133} & 4.139 & 4.134   \\ 
        User Q3 & 4.038 & 4.032 & \underline{4.026}   \\ 
        User Q4 & \underline{3.457} & 3.472 & 3.465   \\ 
        \hline
      \end{tabular}
    \end{center}
  \fi

\end{table*}


